---
title: "Fantasy Football Team Report"
author: "Dusty Turner"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: default
params:
  team: '''R''m Chair_Quarterback'
  week: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, error=FALSE, message=FALSE)
library(tidyverse)
library(kableExtra)
library(DT)
options(DT.options = list(pageLength = 100, language = list(search = 'Filter:')))
```

# Fantasy Football Report for `r params$team`: 
# Week `r params$week`

```{r, echo=FALSE}
ffdata = read_csv("FantasyFootballData.csv")
leaguedata = read_csv("weekbyweekresults.csv")
leaguedatasimple = read_csv("weekbyweekresultssimple.csv")
WeeklyEstimates = 
  ffdata %>% filter(!is.na(Team))
```

```{r}
ffdata %>%
    # filter(scoringPeriodId==4) %>%
  filter(scoringPeriodId==params$week) %>%
  filter(Starter=="Starter") %>%
  select(Actual,Predicted,Player,ActualMinusPredicted) %>%
  summarise(mean=mean(Actual, na.rm = TRUE))  -> weekleaguemean

ffdata %>%
  # filter(Team=="'R'm Chair_Quarterback") %>%
  # filter(scoringPeriodId==1) %>%
  filter(Team==params$team) %>%
  filter(scoringPeriodId==params$week) %>%
  filter(Starter=="Starter") %>%
  summarise(sum = sum(Actual, na.rm = TRUE)) -> weakteamsum

ffdata %>%
  # filter(scoringPeriodId==3) %>%
  filter(scoringPeriodId==params$week) %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(Team)) %>%
  group_by(Team) %>%
  summarise(score = sum(Actual, na.rm = TRUE)) %>%
  filter(score==max(score)|score==min(score)) -> topandbottom
```

### Week Summary

1. This week, your team scored `r weakteamsum$sum`; compared to the league average of `r round(weekleaguemean$mean*9,2)`.

2. The Mug winner is `r topandbottom$Team[1]`.  `r ifelse(params$team==topandbottom$Team[1], "Congratulations on winning the Mug!","Sorry you didn't win the mug :(.")
`

3. The Plunger winner is `r topandbottom$Team[2]`. `r ifelse(params$team==topandbottom$Team[2], "Congradulations on winning the Plunger!","Congradulations on avoiding the Plunger :)." )`

### Did Your Team Measure Up To Expectations?

```{r,fig.align='center'}
ffdata %>%
  # filter(Team=="'R'm Chair_Quarterback") %>%
  # filter(scoringPeriodId==3) %>%
  filter(Team==params$team) %>%
  filter(scoringPeriodId==params$week) %>%
  filter(Starter=="Starter") %>%
  select(Actual,Predicted,Player,ActualMinusPredicted) %>%
  mutate(order = rank(Predicted)) %>%
  mutate(Performance = if_else(Actual>Predicted,"Overachiever","Underachiever")) %>%
  ggplot(aes(group = Player)) +
  geom_point(aes(x=Actual,y=1), size = 1) +
  geom_point(aes(x=Predicted,y=1), size = 5) +
  geom_segment(aes(x=Predicted,xend=Actual,y=1,yend=1,color = Performance), arrow = arrow()) +
  # geom_vline(aes(xintercept = weekleaguemean$mean)) +
  # facet_grid(forcats::fct_reorder(Player,order)~.) +
  facet_wrap(~forcats::fct_reorder(Player,order), ncol = 1) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())+
        # legend.position = "none")  +
  scale_color_manual(values = c("green","red")) +
  labs(y="", x = "Points: Predicted (Big Dot) and Actual (Small Dot) Performance", 
       # title = "Did Your Team Over or Under Achieve?",
       caption = "Black line denoting the league average score is included for reference")

```

### Which Players Measured Up to Expectations?

```{r}
ffdata %>%
  filter(Team==params$team) %>%
  filter(scoringPeriodId==params$week) %>%
  mutate(Predicted = round(Predicted,2)) %>%
  mutate(ActualMinusPredicted = round(ActualMinusPredicted,2)) %>%
  arrange(desc(Starter),playerrosterslot) %>%
  select(Player,Position,SlottedPosition,Starter,Actual,Predicted,ActualMinusPredicted) %>%
  DT::datatable(escape = FALSE,options = list(lengthChange = FALSE, dom = 't')) %>%
  formatStyle('ActualMinusPredicted', backgroundColor = styleInterval(c(-100,-10,-5,0,5,10,100),  c("black","red","lightpink","khaki","khaki","lightgreen","green","black")))
```

### Optimal Lineup

How close were you to achieving your optimal lineup?

```{r include=FALSE}
ranked = 
ffdata %>%
  select(scoringPeriodId,Player,Position,Team,Actual) %>%
  group_by(Team,Position,scoringPeriodId) %>%
  mutate(Rank = rank(-Actual, ties.method = "first")) %>%
  arrange(Team) %>%
  filter(!is.na(Team))

rankedqbs =
ranked %>%
  filter(Position == "Quarterback") %>%
  filter(Rank == 1) %>%
  mutate(PosOrder = 1)

rankedrbs =
ranked %>%
  filter(Position == "Running Back") %>%
  filter(Rank <= 2) %>%
  mutate(PosOrder = 2)

rankedwrs =
ranked %>%
  filter(Position == "Wide Receiver") %>%
  filter(Rank <= 2) %>%
  mutate(PosOrder = 3)

rankedtes =
ranked %>%
  filter(Position == "Tight End") %>%
  filter(Rank <= 1) %>%
  mutate(PosOrder = 4)

rankedk =
ranked %>%
  filter(Position == "Kicker") %>%
  filter(Rank <= 1) %>%
  mutate(PosOrder = 6)

rankeddef =
ranked %>%
  filter(Position == "Defense") %>%
  filter(Rank <= 1) %>%
  mutate(PosOrder = 7)

rankedfl =
ranked %>%
  anti_join(
    bind_rows(rankedrbs,rankedwrs,rankedtes)
  ) %>%
  filter(Position %in% c("Running Back", "Wide Receiver", "Tight End")) %>%
  group_by(scoringPeriodId,Team) %>%
  mutate(Rank = rank(-Actual, ties.method = "first")) %>%
  filter(Rank==1) %>%
  mutate(Position="Flex") %>%
  mutate(PosOrder = 5)

optimallineup = 
bind_rows(rankedqbs,rankedrbs,rankedwrs,rankedtes,rankedfl,rankeddef,rankedk) %>%
  arrange(Team,scoringPeriodId,PosOrder)
```


```{r}
WeeklyEstimates1 = 
WeeklyEstimates %>%
  left_join(
    tibble(SlottedPosition = c("Quarterback", "Running Back","Wide Receiver","Tight End", "Flex", "Kicker", "Defense"),
           PosOrder = 1:7)
  )

data =
optimallineup %>% 
  mutate(Lineup="Best") %>%
  bind_rows(

  WeeklyEstimates1 %>% as_tibble() %>%
  filter(Starter=="Starter") %>%
  select(scoringPeriodId,Player,SlottedPosition,Team,Actual,PosOrder) %>%
  mutate(Lineup="Actual") %>%
  rename(Position = SlottedPosition) %>%
  arrange(Team,scoringPeriodId,PosOrder)
) %>%
  filter(Team==params$team) %>%
  # filter(Team=="'R'm Chair_Quarterback") %>%
  rename(Week = scoringPeriodId) %>%
  ungroup()

data %>%
  filter(Lineup=="Best") %>%
  rename(BestScored = Actual, BestPlayer = Player) %>%
  select(Week,Position,BestPlayer,BestScored) %>%
  bind_cols(
    data %>%
  filter(Lineup=="Actual") %>%
  rename(ActualScored = Actual, ActualPlayer = Player) %>%
  select(ActualPlayer,ActualScored)
  ) %>%
  mutate(Difference = round(BestScored-ActualScored,2)) %>%
  arrange(desc(Week)) %>%
  DT::datatable(options = list(lengthMenu = c(9,18,27),pageLength = 9))

# data %>%
#   ggplot(aes(x=Actual,y=MissedBestBy, color = MissedBestBy)) +
#   geom_point() + 
#   # geom_abline() +
#   ggrepel::geom_text_repel(aes(label = paste0(Team,": " ,round(Best,0))), max.iter = 100000) +
#   theme(legend.position = "none") +
#   # xlim(limmin,limmax) + 
#   ylim(0,limmax)+
#   labs(x="Actual Points Scored", y = "Points Short of Optimal Line Up", 
#        title = "How Good of a Coach Are You?",
#        subtitle = "Closer to the x axis means better coach",
#        caption = "Low and right means good player selection and high points scored.")
```

