---
title: "Fantasy Football Team Report"
author: "Dusty Turner"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
params: 
  team: "'R'm Chair_Quarterback"
  week: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, error=FALSE, message=FALSE)
library(tidyverse)
library(kableExtra)
library(DT)
options(DT.options = list(pageLength = 100, language = list(search = 'Filter:')))
```

# Fantasy Football Report for `r params$team`: Week `r params$week`

```{r, echo=FALSE}
ffdata = read_csv("FantasyFootballData.csv")
```

```{r}
ffdata %>%
    # filter(scoringPeriodId==1) %>%
  filter(scoringPeriodId==params$week) %>%
  filter(Starter=="Starter") %>%
  select(Actual,Predicted,Player,ActualMinusPredicted) %>%
  summarise(mean=mean(Actual))  -> weekleaguemean

ffdata %>%
  # filter(Team=="'R'm Chair_Quarterback") %>%
  # filter(scoringPeriodId==1) %>%
  filter(Team==params$team) %>%
  filter(scoringPeriodId==params$week) %>%
  filter(Starter=="Starter") %>%
  summarise(sum = sum(Actual)) -> weakteamsum

ffdata %>%
  # filter(scoringPeriodId==1) %>%
  filter(scoringPeriodId==params$week) %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(Team)) %>%
  group_by(Team) %>%
  summarise(score = sum(Actual)) %>%
  filter(score==max(score)|score==min(score)) -> topandbottom

# ifelse(params$team==topandbottom$Team[1], "Congradulations on winning the Plunger","Congradulations on avoiding the Plunger" )
```

### Week Summary

1. This week, your team scored `r weakteamsum$sum`; compared to the league average of `r round(weekleaguemean$mean*9,2)`.

2. The Plunger winner is `r topandbottom$Team[1]`.  `r ifelse(params$team==topandbottom$Team[1], "Congradulations on winning the Plunger","Congradulations on avoiding the Plunger" )`!
`

3. The Mug winner is `r topandbottom$Team[2]`. `r ifelse(params$team==topandbottom$Team[2], "Congradulations on winning the Mug!","Sorry, you didn't win the mug." )`

### Visually

```{r}
ffdata %>%
  # filter(Team=="'R'm Chair_Quarterback") %>%
  # filter(scoringPeriodId==1) %>%
  filter(Team==params$team) %>%
  filter(scoringPeriodId==params$week) %>%
  filter(Starter=="Starter") %>%
  select(Actual,Predicted,Player,ActualMinusPredicted) %>%
  mutate(order = rank(Actual)) %>%
  gather(Type,Points,-Player,-ActualMinusPredicted,-order) %>%
  ggplot(aes(x=Points,y=1, group = Player)) +
  geom_point(aes(color = Type, size = Type)) +
  geom_line(aes(color = Type)) +
  geom_vline(aes(xintercept = weekleaguemean$mean)) +
  # facet_grid(forcats::fct_reorder(Player,order)~.) +
  facet_wrap(~forcats::fct_reorder(Player,order), ncol = 1) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())  +
  scale_color_manual(values = c("blue","tan")) +
  labs(y="", x = "Points: Predicted and Actual", title = "Did Your Team Over or Under Achieve?",
       subtitle = "Blue Dots on the left are Underachievers", caption = "Black line denoting the league average score is included for refrence")

```


```{r}
ffdata %>%
  filter(Team==params$team) %>%
  filter(scoringPeriodId==params$week) %>%
  mutate(Predicted = round(Predicted,2)) %>%
  mutate(ActualMinusPredicted = round(ActualMinusPredicted,2)) %>%
  arrange(desc(Starter),playerrosterslot) %>%
  select(Player,Position,SlottedPosition,Starter,Actual,Predicted,ActualMinusPredicted) %>%
  DT::datatable(escape = FALSE) %>%
  formatStyle('ActualMinusPredicted', backgroundColor = styleInterval(c(-100,-10,-5,0,5,10,100),  c("black","red","lightpink","khaki","khaki","lightgreen","green","black")))
```


