---
title: "Fantasy Football League Report"
author: "Dusty Turner"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE)
options(knitr.table.format = "html")
library(tidyverse)
library(kableExtra)
library(formattable)
playerdata = read_csv("FantasyFootballData.csv")
leaguedata = read_csv("weekbyweekresults.csv")
leaguedatasimple = read_csv("weekbyweekresultssimple.csv")
week = leaguedatasimple %>% select(awaypoints) %>% as.data.frame() %>%
  filter(awaypoints!=0) %>% nrow()/5

```

# League Update

Hello all, welcome to OA Division Fantasy Football Week `r week` Update!

## Mug (and Plunger) Standings

```{r}

teams = 
leaguedatasimple %>%
  select(AwayTeam,HomeTeam)%>%
  gather(direction, Team) %>%
  select(Team) %>%
  distinct()

currentstandings = 
leaguedatasimple %>%
  filter(homepoints!=0) %>%
  group_by(winner) %>%
  mutate(Wins = n()) %>%
  select(winner, Wins) %>%
  distinct() %>%
  rename(Team = winner) %>%
  right_join(teams) %>%
  mutate(Wins = replace_na(Wins, 0)) %>%
  mutate(Losses = week - Wins) %>%
  mutate(WinPerc =Wins/(Wins+Losses))

weeksummedstandings =
leaguedata %>%
  filter(weekID <=week) %>%
  mutate(Team = if_else(Location=="awaypoints",AwayTeam,HomeTeam)) %>%
  group_by(Team) %>%
  summarise(EveryWeekWins = sum(EWETW), EveryWeekLosses = sum(EWETL)) %>%
  mutate(EveryWeekPerc = EveryWeekWins/(EveryWeekWins+EveryWeekLosses))

standings = 
  currentstandings %>%
  left_join(weeksummedstandings) 

leaguedatasimple %>%
  filter(homepoints!=0) %>%
  group_by(winner) %>%
  mutate(Wins = n()) %>% ungroup() %>%
  select(awaypoints,homepoints,AwayTeam,HomeTeam,weekID) %>%
  gather(Location,Team,-weekID,-awaypoints,-homepoints) %>%
  mutate(points = if_else(Location=="AwayTeam",awaypoints,homepoints)) %>% 
  select(Team,points,weekID) %>%
  group_by(weekID) %>%
  filter(points == min(points) | points == max(points)) %>%
  mutate(Award = if_else(points==max(points),"Mug","Plunger")) %>%
  ungroup() %>%
  count(Award, Team)  -> awards

leaguedatasimple %>%
  filter(homepoints!=0) %>%
  filter(weekID == max(weekID)) %>%
  select(awaypoints,homepoints,AwayTeam,HomeTeam,weekID) %>%
  gather(Location,Team,-weekID,-awaypoints,-homepoints) %>%
  mutate(points = if_else(Location=="AwayTeam",awaypoints,homepoints)) %>% 
  select(Team,points,weekID) %>%
  filter(points == min(points) | points == max(points)) %>%
  mutate(Award = if_else(points==max(points),"Mug","Plunger")) %>%
  ungroup() %>%
  count(Award, Team) %>%
  arrange(Award) -> awardsthisweek

```

Everyone congratulate `r awardsthisweek$Team[1]` for winning the mug.

Hardy cajoling goes out to `r awardsthisweek$Team[2]` for earning the plunger this week.  Hats off for your hard work1

Here is the current Mug tally!

```{r}
awards %>%
  filter(Award=="Mug") %>% 
  kable("html", escape = F) %>%
  kable_styling("hover", full_width = F)
```

Here is the current Plunger tally!

```{r}
awards %>%
  filter(Award=="Plunger") %>% 
  kable("html", escape = F) %>%
  kable_styling("hover", full_width = F)
```


## Standings

The standings are current through week `r week`.

Every week wins are how many wins your team would have if you played every team every week.  Every week losses and every week win percentage are calculated accordingly.  This is meant to be a measure of how 'lucky' your team is. 

```{r}
standings %>%
  mutate(EveryWeekPerc = round(EveryWeekPerc,2)) %>%
  arrange(desc(WinPerc)) %>%
      kable("html", escape = F) %>%
  kable_styling("hover", full_width = F) 

```


A graphical view of the above information.  Low and right are lucky. Up and left are unlucky.

```{r}
standings %>%
  ggplot(aes(x=WinPerc,y=EveryWeekPerc, color = WinPerc)) +
  geom_point() +
  geom_abline() +
  annotate("text", x = .9, y = .05, label = "Lucky", color = "green") +
  annotate("text", x = .1, y = .95, label = "Unlucky", color = "darkgoldenrod") +
  annotate("text", x = .9, y = .95, label = "Good", color = "blue") +
  annotate("text", x = .1, y = .05, label = "Bad", color = "red") +
  ggrepel::geom_text_repel(aes(label = Team)) +
  labs(x="Traditional Winning Percentage", y = "Every Team Every Week Win Percentage",
       title = "Are You As Good (or bad) As You Should Be?",
       caption = "The lighter the blue the higher the traditional win percentage.") +
  theme(legend.position = "none")
  
```


### What Teams Are Living Up to Expectations This Week?

```{r}
WeeklyEstimates = 
  playerdata %>% filter(!is.na(Team))
```


```{r}
WeeklyEstimates %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(ActualMinusPredicted)) %>% 
  filter(scoringPeriodId==week) %>% 
  group_by(Team, scoringPeriodId, Starter) %>% 
  summarise(TeamActual = sum(Actual),
            TeamPredicted = sum(Predicted)) %>%
  ungroup() %>%
  select(TeamActual,TeamPredicted) %>%
  gather(Team, Points) %>%
  summarise(max = max(Points), min = min(Points)) -> limvec

WeeklyEstimates %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(ActualMinusPredicted)) %>%
  filter(scoringPeriodId==week) %>%
  group_by(Team, scoringPeriodId, Starter) %>% 
  # filter(Team == "'R'm Chair_Quarterback") %>% arrange(Starter) %>%
  summarise(TeamActual = sum(Actual),
            TeamPredicted = sum(Predicted)) %>%
  ungroup() %>%
  mutate(Team = str_replace(Team,"_"," ")) %>%
  mutate(TeamActualMinusPredicted = TeamActual-TeamPredicted) %>%
  ggplot(aes(x=TeamPredicted,y=TeamActual, color = TeamActualMinusPredicted)) +
  geom_point() +
  geom_abline() +
  xlim(limvec$min,limvec$max) +   ylim(limvec$min,limvec$max) +
  ggrepel::geom_text_repel(aes(label = Team), max.iter = 10000) +
  labs(x="Predicted Points Scored", y="Actual Points Scored", title = paste0("Projected vs Actual Points Scored in Week ", week), color = "Actual-Predicted") 
```

### What Teams Are Living Up to Expectations All Weeks?


```{r}
WeeklyEstimates %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(ActualMinusPredicted)) %>%
  # filter(scoringPeriodId==week) %>%
  group_by(Team, Starter) %>% 
  summarise(TeamActual = sum(Actual),
            TeamPredicted = sum(Predicted)) %>%
  ungroup() %>%
  select(TeamActual,TeamPredicted) %>%
  gather(Team, Points) %>%
  summarise(max = max(Points), min = min(Points)) -> limvec

WeeklyEstimates %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(ActualMinusPredicted)) %>%
  # filter(scoringPeriodId==week) %>%
  group_by(Team, Starter) %>% 
  # filter(Team == "'R'm Chair_Quarterback") %>% arrange(Starter) %>%
  summarise(TeamActual = sum(Actual),
            TeamPredicted = sum(Predicted)) %>%
  mutate(TeamActualMinusPredicted = TeamActual-TeamPredicted) %>%
  ungroup() %>%
  mutate(Team = str_replace(Team,"_"," ")) %>%
  ggplot(aes(x=TeamPredicted,y=TeamActual, color = TeamActualMinusPredicted)) +
  geom_point() +
  geom_abline() +
  xlim(limvec$min,limvec$max) +   ylim(limvec$min,limvec$max) +
  ggrepel::geom_text_repel(aes(label = Team), max.iter = 10000) +
  labs(x="Predicted Points Scored", y="Actual Points Scored", title = paste0("Projected vs Actual Points Scored All Season"), color = "Actual-Predicted") 
```



### Whose Team's Starters Are Meeting Expectations?

This table shows the Team's actual points scored minus their projected points scored in week `r week`.

```{r echo=FALSE}

## expectation performance by most recent week
WeeklyEstimates %>%
  filter(!is.na(Team)) %>%
  arrange(-ActualMinusPredicted) %>%
  group_by(Team, scoringPeriodId) %>%
  summarise(MeanActualMinusPredicted = mean(ActualMinusPredicted, na.rm = TRUE)) %>%
  arrange(MeanActualMinusPredicted) %>%
  filter(scoringPeriodId==week) %>%
  select(-scoringPeriodId) %>% 
  rename(Delta = MeanActualMinusPredicted) %>%
  mutate(
      Delta = round(Delta,2),
      Delta = ifelse(Delta > 0,
                  cell_spec(Delta, "html", color = "green", bold = T),
                  cell_spec(Delta, "html", color = "red", italic = T))
    ) %>%
    kable("html", escape = F) %>%
  kable_styling("hover", full_width = F) 
```

This table show's how each team has actually performed vs their projection (on average) all season.

```{r}
## expectation performance overall
WeeklyEstimates %>%
  filter(!is.na(Team)) %>%
  arrange(-ActualMinusPredicted) %>%
  group_by(Team) %>%
  summarise(MeanActualMinusPredicted = mean(ActualMinusPredicted, na.rm = TRUE)) %>%
  arrange(MeanActualMinusPredicted) %>%
    rename(Delta = MeanActualMinusPredicted) %>%
    mutate(
      Delta = round(Delta,2),
      Delta = ifelse(Delta > 0,
                  cell_spec(Delta, "html", color = "green", bold = T),
                  cell_spec(Delta, "html", color = "red", italic = T))
    ) %>%
    kable("html", escape = F) %>%
  kable_styling("hover", full_width = F) 
```

## Optimal Lineup

Who is the best coach?  How much have you missed your optimal line up this week.  The plot below reflects how many points you scored vs your best line up.

```{r include=FALSE}
ranked = 
playerdata %>%
  select(scoringPeriodId,Player,Position,Team,Actual) %>%
  group_by(Team,Position,scoringPeriodId) %>%
  mutate(Rank = rank(-Actual, ties.method = "first")) %>%
  arrange(Team) %>%
  filter(!is.na(Team))

rankedqbs =
ranked %>%
  filter(Position == "Quarterback") %>%
  filter(Rank == 1) %>%
  mutate(PosOrder = 1)

rankedrbs =
ranked %>%
  filter(Position == "Running Back") %>%
  filter(Rank <= 2) %>%
  mutate(PosOrder = 2)

rankedwrs =
ranked %>%
  filter(Position == "Wide Receiver") %>%
  filter(Rank <= 2) %>%
  mutate(PosOrder = 3)

rankedtes =
ranked %>%
  filter(Position == "Tight End") %>%
  filter(Rank <= 1) %>%
  mutate(PosOrder = 4)

rankedk =
ranked %>%
  filter(Position == "Kicker") %>%
  filter(Rank <= 1) %>%
  mutate(PosOrder = 6)

rankeddef =
ranked %>%
  filter(Position == "Defense") %>%
  filter(Rank <= 1) %>%
  mutate(PosOrder = 7)

rankedfl =
ranked %>%
  anti_join(
    bind_rows(rankedrbs,rankedwrs,rankedtes)
  ) %>%
  filter(Position %in% c("Running Back", "Wide Receiver", "Tight End")) %>%
  group_by(scoringPeriodId,Team) %>%
  mutate(Rank = rank(-Actual, ties.method = "first")) %>%
  filter(Rank==1) %>%
  mutate(Position="Flex") %>%
  mutate(PosOrder = 5)

optimallineup = 
bind_rows(rankedqbs,rankedrbs,rankedwrs,rankedtes,rankedfl,rankeddef,rankedk) %>%
  arrange(Team,scoringPeriodId,PosOrder)
```


```{r}
optimallineup %>% 
  mutate(Lineup="Best") %>%
  bind_rows(

WeeklyEstimates %>% as_tibble() %>%
  filter(Starter=="Starter") %>%
  select(scoringPeriodId,Player,Position,Team,Actual) %>%
  mutate(Lineup="Actual")
) %>%
  group_by(Team,Lineup) %>%
  summarise(Points = sum(Actual)) %>% 
  spread(Lineup,Points) %>% 
  ungroup() %>%
  mutate(Team = str_replace(Team,"_"," ")) %>%
  filter(!is.na(Team)) %>%
  mutate(MissedBestBy = Best-Actual) -> data 

limmax = max(data$MissedBestBy)
# limmin = min(data$Actual,data$Best)

data %>%
  ggplot(aes(x=Actual,y=MissedBestBy, color = MissedBestBy)) +
  geom_point() + 
  # geom_abline() +
  ggrepel::geom_text_repel(aes(label = paste0(Team,": " ,round(Best,0))), max.iter = 100000) +
  theme(legend.position = "none") +
  # xlim(limmin,limmax) + 
  ylim(0,limmax)+
  labs(x="Actual Points Scored", y = "Points Short of Optimal Line Up", 
       title = "How Good of a Coach Are You?",
       subtitle = "Closer to the x axis means better coach",
       caption = "Low and right means good player selection and high points scored.")
```

How good of a coach were you in week `r week`?

```{r}
optimallineup %>% 
  mutate(Lineup="Best") %>%
  bind_rows(

WeeklyEstimates %>% as_tibble() %>%
  filter(Starter=="Starter") %>%
  select(scoringPeriodId,Player,Position,Team,Actual) %>%
  mutate(Lineup="Actual")
) %>% 
  filter(scoringPeriodId == week) %>%
  group_by(Team,Lineup) %>%
  summarise(Points = sum(Actual)) %>% 
  spread(Lineup,Points) %>% 
  ungroup() %>%
  mutate(Team = str_replace(Team,"_"," ")) %>%
  filter(!is.na(Team)) %>%
  mutate(MissedBestBy = Best-Actual) -> data 

limmax = max(data$MissedBestBy)


data %>%
  filter(week == week) %>%
  mutate(MissedBestBy = Best-Actual) %>%
  ggplot(aes(x=Actual,y=MissedBestBy, color = MissedBestBy)) +
  geom_point() + 
  # geom_abline() +
  ggrepel::geom_text_repel(aes(label = paste0(Team,": " ,round(Best,0))), max.iter = 100000) +
  theme(legend.position = "none") +
  # xlim(limmin,limmax) + 
  ylim(0,limmax)+
  labs(x="Actual Points Scored", y = "Points Short of Optimal Line Up", 
       title = "How Good of a Coach Are You?",
       subtitle = "Closer to the x axis means better coach",
       caption = "Low and right means good player selection and high points scored.")

```



## A Look Across the League: 

### What Positions Are Predicted Most Accurately

```{r}
WeeklyEstimates %>% as_tibble() %>% 
  filter(!is.na(ActualMinusPredicted)) %>%
  group_by(Position) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  ggplot(aes(x=ActualMinusPredicted)) +
  geom_density(aes(fill = Position)) +
  geom_histogram(aes(fill = Position)) +
  facet_wrap(~forcats::fct_reorder(Position,-n), ncol = 1) +
  geom_vline(aes(xintercept = 0, color = "red")) +
  theme(legend.position = "none", 
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  labs(title = "Histogram of How Bad (or Good) ESPN Projects Each Position", 
       y = "", x = "Actual Points Scored Minus Predicted")
```



<!-- ```{r} -->

<!-- WeeklyEstimates %>% -->
<!--   filter(!is.na(ActualMinusPredicted)) %>% -->
<!--   group_by(Team, scoringPeriodId, Starter)  -->


<!-- WeeklyEstimates %>% as_tibble() %>%  -->
<!--   filter(!is.na(ActualMinusPredicted)) %>% -->
<!--   ggplot(aes(x=ActualMinusPredicted)) + -->
<!--   geom_histogram() -->

<!-- WeeklyEstimates %>% as_tibble() %>%  -->
<!--   filter(!is.na(ActualMinusPredicted)) %>% -->
<!--   ggplot(aes(x=Predicted,y=Actual)) + -->
<!--   geom_point() + -->
<!--   geom_smooth(method = "lm") -->


<!-- WeeklyEstimates %>% as_tibble() %>%  -->
<!--   filter(!is.na(ActualMinusPredicted)) %>% -->
<!--   ggplot(aes(x=Predicted,y=Actual, color = Team)) + -->
<!--   geom_point() + -->
<!--   # facet_grid(Starter~Team) + -->
<!--   # facet_wrap(~Team, ncol = 2) + -->
<!--   geom_abline(aes(slope =1, intercept = 0),color = "red") -->
<!--   # theme(legend.position = "none") -->

<!-- ``` -->
