---
title: "League Report Dashboard"
author: "Dusty Turner"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  flexdashboard::flex_dashboard:
    theme: spacelab
    vertical_layout: scroll
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE)
options(knitr.table.format = "html")
library(tidyverse)
library(kableExtra)
library(formattable)
playerdata = read_csv("FantasyFootballData.csv")
leaguedata = read_csv("weekbyweekresults.csv")
leaguedatasimple = read_csv("weekbyweekresultssimple.csv")
week = leaguedatasimple %>% select(awaypoints) %>% as.data.frame() %>%
  filter(awaypoints!=0) %>% nrow()/5

```


Weekly Update
=====================================  
    
### Mug (and Plunger) Standings

Hello all, welcome to OA Division Fantasy Football Week `r week` Update!
    
```{r}

teams = 
leaguedatasimple %>%
  select(AwayTeam,HomeTeam)%>%
  gather(direction, Team) %>%
  select(Team) %>%
  distinct()

currentstandings = 
leaguedatasimple %>%
  filter(homepoints!=0) %>%
  group_by(winner) %>%
  mutate(Wins = n()) %>%
  select(winner, Wins) %>%
  distinct() %>%
  rename(Team = winner) %>%
  right_join(teams) %>%
  mutate(Wins = replace_na(Wins, 0)) %>%
  mutate(Losses = week - Wins) %>%
  mutate(WinPerc =Wins/(Wins+Losses))

weeksummedstandings =
leaguedata %>%
  filter(weekID <=week) %>%
  mutate(Team = if_else(Location=="awaypoints",AwayTeam,HomeTeam)) %>%
  group_by(Team) %>%
  summarise(EveryWeekWins = sum(EWETW), EveryWeekLosses = sum(EWETL)) %>%
  mutate(EveryWeekPerc = EveryWeekWins/(EveryWeekWins+EveryWeekLosses))

standings = 
  currentstandings %>%
  left_join(weeksummedstandings) 

leaguedatasimple %>%
  filter(homepoints!=0) %>%
  group_by(winner) %>%
  mutate(Wins = n()) %>% ungroup() %>%
  select(awaypoints,homepoints,AwayTeam,HomeTeam,weekID) %>%
  gather(Location,Team,-weekID,-awaypoints,-homepoints) %>%
  mutate(points = if_else(Location=="AwayTeam",awaypoints,homepoints)) %>% 
  select(Team,points,weekID) %>%
  group_by(weekID) %>%
  filter(points == min(points) | points == max(points)) %>%
  mutate(Award = if_else(points==max(points),"Mug","Plunger")) %>%
  ungroup() %>%
  count(Award, Team)  -> awards

leaguedatasimple %>%
  filter(homepoints!=0) %>%
  filter(weekID == max(weekID)) %>%
  select(awaypoints,homepoints,AwayTeam,HomeTeam,weekID) %>%
  gather(Location,Team,-weekID,-awaypoints,-homepoints) %>%
  mutate(points = if_else(Location=="AwayTeam",awaypoints,homepoints)) %>% 
  select(Team,points,weekID) %>%
  filter(points == min(points) | points == max(points)) %>%
  mutate(Award = if_else(points==max(points),"Mug","Plunger")) %>%
  ungroup() %>%
  count(Award, Team) %>%
  arrange(Award) -> awardsthisweek

```

Mug: Everyone congratulate `r awardsthisweek$Team[1]` for winning the mug.

Plunger: Hardy cajoling goes out to `r awardsthisweek$Team[2]` for earning the plunger this week.  Hats off for your hard work!

### Current Mug tally:

```{r}
awards %>%
  filter(Award=="Mug") %>% 
  arrange(-n) %>%
  kable("html", escape = F) %>%
  kable_styling("hover", full_width = F)
```

### Current Plunger tally:

```{r}
awards %>%
  filter(Award=="Plunger") %>% 
  arrange(-n) %>%
  kable("html", escape = F) %>%
  kable_styling("hover", full_width = F)
```
 


Standings {data-orientation=columns}
=====================================     


Column {.tabset}
-------------------------------------
   
### Standings
    

```{r}
standings %>%
  mutate(EveryWeekPerc = round(EveryWeekPerc,2)) %>%
  arrange(desc(WinPerc)) %>%
  mutate(WinPerc = round(WinPerc,2)) %>%
  DT::datatable(escape = FALSE,options = list(lengthChange = FALSE, dom = 't', list(pageLength = 10))) 
  # kable("html", escape = F) %>%
  # kable_styling("hover", full_width = F) 

```

### Graphic

```{r}
standings %>% ungroup() %>%
  mutate(dist = ((-1*WinPerc)+(1*EveryWeekPerc+0))/sqrt(1+1)) %>%
  # mutate(mult = (dist*1000)) %>%
  mutate(rank = rank(dist)) %>%
  ggplot(aes(x=WinPerc,y=EveryWeekPerc, color = dist)) +
  geom_point() +
  geom_abline() +
  annotate("text", x = .9, y = .05, label = "Lucky", color = "green") +
  annotate("text", x = .1, y = .95, label = "Unlucky", color = "darkgoldenrod") +
  annotate("text", x = .9, y = .95, label = "Good", color = "blue") +
  annotate("text", x = .1, y = .05, label = "Bad", color = "red") +
  ggrepel::geom_text_repel(aes(label = Team)) +
  scale_color_gradient2(low = "dark green", mid = "black", high = "red", midpoint = 0) +
  labs(x="Traditional Winning Percentage", y = "Every Team Every Week Win Percentage",
       title = "Are You As Good (or bad) As Your Record Indicates?",
       caption = "Green = Lucky \n Red = Unlucky") + 
  theme(legend.position = "none")
  
```



Column
-------------------------------------

### Explanation

The standings are current through week `r week`.

Every week wins are how many wins your team would have if you played every team every week.  

Every week losses and every week win percentage are calculated accordingly.  

This is meant to be a measure of how 'lucky' your team is. 

A graphical view of the above information.  Low and right are lucky. Up and left are unlucky.
.  
.  
.  
.  
.  
.  
.  
.  
.  
.  
.  
.  
.  
.  
.  
.  
.  
.  

Expectations
=====================================  

Row
-------------------------------------
    
### What Teams Are Living Up to Expectations This Week?

```{r}
WeeklyEstimates = 
  playerdata %>% filter(!is.na(Team))
```


```{r}
WeeklyEstimates %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(ActualMinusPredicted)) %>% 
  filter(scoringPeriodId==week) %>% 
  group_by(Team, scoringPeriodId, Starter) %>% 
  summarise(TeamActual = sum(Actual),
            TeamPredicted = sum(Predicted)) %>%
  ungroup() %>%
  select(TeamActual,TeamPredicted) %>%
  gather(Team, Points) %>%
  summarise(max = max(Points), min = min(Points)) -> limvec

WeeklyEstimates %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(ActualMinusPredicted)) %>%
  filter(scoringPeriodId==week) %>%
  group_by(Team, scoringPeriodId, Starter) %>% 
  # filter(Team == "'R'm Chair_Quarterback") %>% arrange(Starter) %>%
  summarise(TeamActual = sum(Actual),
            TeamPredicted = sum(Predicted)) %>%
  ungroup() %>%
  mutate(Team = str_replace(Team,"_"," ")) %>%
  mutate(TeamActualMinusPredicted = TeamActual-TeamPredicted) %>%
  mutate(dist = ((-1*TeamPredicted)+(1*TeamActual+0))/sqrt(1+1)) %>%
  ggplot(aes(x=TeamPredicted,y=TeamActual, color = TeamActualMinusPredicted)) +
  geom_point() +
  geom_abline() +
  xlim(limvec$min,limvec$max) +   ylim(limvec$min,limvec$max) +
  ggrepel::geom_text_repel(aes(label = Team), max.iter = 10000) +
  scale_color_gradient2(low = "red", mid = "black", high = "dark green", midpoint = 0) +
  labs(x="Predicted Points Scored", 
       y="Actual Points Scored", 
       title = paste0("Projected vs Actual Points Scored in Week ", week), 
       color = "Actual-Predicted",
       caption = "Green = Lucky \n Red = Unlucky") 
```
 
### Whose Team's Starters Are Meeting Expectations?

This table shows the Team's starter's actual points scored minus their projected points scored in week `r week`.

```{r echo=FALSE}

## expectation performance by most recent week
WeeklyEstimates %>%
  filter(!is.na(Team)) %>%
  filter(Starter == "Starter") %>%
  arrange(-ActualMinusPredicted) %>%
  group_by(Team, scoringPeriodId) %>%
  summarise(totalActualMinusPredicted = sum(ActualMinusPredicted, na.rm = TRUE)) %>%
  arrange(totalActualMinusPredicted) %>%
  filter(scoringPeriodId==week) %>%
  select(-scoringPeriodId) %>% 
  rename(Delta = totalActualMinusPredicted) %>%
  mutate(
      Delta = round(Delta,2),
      Delta = ifelse(Delta > 0,
                  cell_spec(Delta, "html", color = "green", bold = T),
                  cell_spec(Delta, "html", color = "red", italic = T))
    ) %>%
    DT::datatable(escape = FALSE,options = list(lengthChange = FALSE, dom = 't')) 

```


Row
-------------------------------------
    
### What Teams Are Living Up to Expectations All Weeks?

```{r}
WeeklyEstimates %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(ActualMinusPredicted)) %>%
  # filter(scoringPeriodId==week) %>%
  group_by(Team, Starter) %>% 
  summarise(TeamActual = sum(Actual),
            TeamPredicted = sum(Predicted)) %>%
  ungroup() %>%
  select(TeamActual,TeamPredicted) %>%
  gather(Team, Points) %>%
  summarise(max = max(Points), min = min(Points)) -> limvec

WeeklyEstimates %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(ActualMinusPredicted)) %>%
  # filter(scoringPeriodId==week) %>%
  group_by(Team, Starter) %>% 
  # filter(Team == "'R'm Chair_Quarterback") %>% arrange(Starter) %>%
  summarise(TeamActual = sum(Actual),
            TeamPredicted = sum(Predicted)) %>%
  mutate(TeamActualMinusPredicted = TeamActual-TeamPredicted) %>%
  ungroup() %>%
  mutate(Team = str_replace(Team,"_"," ")) %>%
  ggplot(aes(x=TeamPredicted,y=TeamActual, color = TeamActualMinusPredicted)) +
  geom_point() +
  geom_abline() +
  xlim(limvec$min,limvec$max) +   ylim(limvec$min,limvec$max) +
  ggrepel::geom_text_repel(aes(label = Team), max.iter = 10000) +
  scale_color_gradient2(low = "red", mid = "black", high = "dark green", midpoint = 0) +
  labs(x="Predicted Points Scored", 
       y="Actual Points Scored", 
       title = paste0("Projected vs Actual Points Scored All Season"), 
       color = "Actual-Predicted",
       caption = "Green = Lucky \n Red = Unlucky")  
```

### Chart 4

This table show's how each team's starters have actually performed vs their projection (on average) all season.

```{r}
## expectation performance overall
WeeklyEstimates %>%
  filter(!is.na(Team)) %>%
  filter(Starter == "Starter") %>%
  arrange(-ActualMinusPredicted) %>%
  group_by(Team) %>%
  summarise(totalActualMinusPredicted = sum(ActualMinusPredicted, na.rm = TRUE)) %>%
  arrange(totalActualMinusPredicted) %>%
    rename(Delta = totalActualMinusPredicted) %>%
    mutate(
      Delta = round(Delta,2),
      Delta = ifelse(Delta > 0,
                  cell_spec(Delta, "html", color = "green", bold = T),
                  cell_spec(Delta, "html", color = "red", italic = T))
    ) %>%
    DT::datatable(escape = FALSE,options = list(lengthChange = FALSE, dom = 't')) 
```

Optimal Lineup
=====================================  
    
Column {.tabset}
-------------------------------------

### Optimal Lineup This Week

Who is the best coach?  

How much have you missed your optimal line up this week?  

The plot below reflects how many 

points you scored vs your best line up.

```{r include=FALSE}
ranked = 
playerdata %>%
  select(scoringPeriodId,Player,Position,Team,Actual) %>%
  group_by(Team,Position,scoringPeriodId) %>%
  mutate(Rank = rank(-Actual, ties.method = "first")) %>%
  arrange(Team) %>%
  filter(!is.na(Team))

rankedqbs =
ranked %>%
  filter(Position == "Quarterback") %>%
  filter(Rank == 1) %>%
  mutate(PosOrder = 1)

rankedrbs =
ranked %>%
  filter(Position == "Running Back") %>%
  filter(Rank <= 2) %>%
  mutate(PosOrder = 2)

rankedwrs =
ranked %>%
  filter(Position == "Wide Receiver") %>%
  filter(Rank <= 2) %>%
  mutate(PosOrder = 3)

rankedtes =
ranked %>%
  filter(Position == "Tight End") %>%
  filter(Rank <= 1) %>%
  mutate(PosOrder = 4)

rankedk =
ranked %>%
  filter(Position == "Kicker") %>%
  filter(Rank <= 1) %>%
  mutate(PosOrder = 6)

rankeddef =
ranked %>%
  filter(Position == "Defense") %>%
  filter(Rank <= 1) %>%
  mutate(PosOrder = 7)

rankedfl =
ranked %>%
  anti_join(
    bind_rows(rankedrbs,rankedwrs,rankedtes)
  ) %>%
  filter(Position %in% c("Running Back", "Wide Receiver", "Tight End")) %>%
  group_by(scoringPeriodId,Team) %>%
  mutate(Rank = rank(-Actual, ties.method = "first")) %>%
  filter(Rank==1) %>%
  mutate(Position="Flex") %>%
  mutate(PosOrder = 5)

optimallineup = 
bind_rows(rankedqbs,rankedrbs,rankedwrs,rankedtes,rankedfl,rankeddef,rankedk) %>%
  arrange(Team,scoringPeriodId,PosOrder)
```



```{r}
helper =
  leaguedatasimple %>%
  mutate(AwayTeam= str_replace(AwayTeam,"_"," ")) %>%
  mutate(HomeTeam= str_replace(HomeTeam,"_"," ")) %>%
  mutate(scoringPeriodId = weekID) %>%
       left_join(

optimallineup %>% 
  ungroup(Team) %>%
  mutate(Lineup="Best") %>%
  mutate(Team= str_replace(Team,"_"," ")) %>%
  mutate(Team = if_else(Team=="Compute  This!", "Compute This!", Team)) %>%
  group_by(Team, scoringPeriodId) %>%
  summarise(Points = sum(Actual)) %>%
  group_by(scoringPeriodId) %>%
  mutate(Rank = rank(-Points)) , by = c("AwayTeam"="Team", "scoringPeriodId" = "scoringPeriodId")) 

BestPointsDataFrame =
helper %>%  
  left_join(
  
  optimallineup %>% 
  ungroup(Team) %>%
  mutate(Lineup="Best") %>%
  mutate(Team= str_replace(Team,"_"," ")) %>%
  mutate(Team = if_else(Team=="Compute  This!", "Compute This!", Team)) %>%
  group_by(Team, scoringPeriodId) %>%
  summarise(Points = sum(Actual)) %>%
  group_by(scoringPeriodId) %>%
  mutate(Rank2 = rank(-Points)) , by = c("HomeTeam"="Team", "scoringPeriodId" = "scoringPeriodId"), suffix = c(".Away",".Home"))

```


```{r}
optimallineup %>% 
  mutate(Lineup="Best") %>%
  bind_rows(

WeeklyEstimates %>% as_tibble() %>%
  filter(Starter=="Starter") %>%
  select(scoringPeriodId,Player,Position,Team,Actual) %>%
  mutate(Lineup="Actual")
) %>%
  group_by(Team,Lineup) %>%
  summarise(Points = sum(Actual)) %>% 
  spread(Lineup,Points) %>% 
  ungroup() %>%
  mutate(Team = str_replace(Team,"_"," ")) %>%
  filter(!is.na(Team)) %>%
  mutate(MissedBestBy = Best-Actual) -> data 

limmax = max(data$MissedBestBy)
# limmin = min(data$Actual,data$Best)

data %>%
  ggplot(aes(x=Actual,y=MissedBestBy, color = MissedBestBy)) +
  geom_point() + 
  # geom_abline() +
  ggrepel::geom_text_repel(aes(label = paste0(Team,": " ,round(Best,0))), max.iter = 100000) +
  theme(legend.position = "none") +
  # xlim(limmin,limmax) + 
  ylim(0,limmax)+
  labs(x="Actual Points Scored", y = "Points Short of Optimal Line Up", 
       title = paste0("How Good of a Coach were You?"),
       subtitle = "Closer to the x axis means better coach",
       caption = "Low and right means good player selection and high points scored.")
```


### Optimial Line Up in Week `r week`

How good of a coach were you in week `r week`?

```{r}
optimallineup %>% 
  mutate(Lineup="Best") %>%
  bind_rows(

WeeklyEstimates %>% as_tibble() %>%
  filter(Starter=="Starter") %>%
  select(scoringPeriodId,Player,Position,Team,Actual) %>%
  mutate(Lineup="Actual")
) %>% 
  filter(scoringPeriodId == week) %>%
  group_by(Team,Lineup) %>%
  summarise(Points = sum(Actual)) %>% 
  spread(Lineup,Points) %>% 
  ungroup() %>%
  mutate(Team = str_replace(Team,"_"," ")) %>%
  filter(!is.na(Team)) %>%
  mutate(MissedBestBy = Best-Actual) -> data 

limmax = max(data$MissedBestBy)


data %>%
  filter(week == week) %>%
  mutate(MissedBestBy = Best-Actual) %>%
  ggplot(aes(x=Actual,y=MissedBestBy, color = MissedBestBy)) +
  geom_point() + 
  # geom_abline() +
  ggrepel::geom_text_repel(aes(label = paste0(Team,": " ,round(Best,0))), max.iter = 100000) +
  theme(legend.position = "none") +
  # xlim(limmin,limmax) + 
  ylim(0,limmax)+
  labs(x="Actual Points Scored", y = "Points Short of Optimal Line Up", 
       title = paste0("How Good of a Coach were you in week ", week, "?"),
       subtitle = "Closer to the x axis means better coach",
       caption = "Low and right means good player selection and high points scored.")

```


### What if you played your best roster each week?

The following table shows how well your team could be doing if you played your best roster every week.


```{r}
BestPointsDataFrame %>%
  filter(homepoints!=0) %>%
  select(Points.Away, AwayTeam,Points.Home,HomeTeam) %>%
  mutate(winner = if_else(Points.Away>Points.Home,AwayTeam,HomeTeam)) %>%
  group_by(winner) %>%
  mutate(Wins = n()) %>%
  select(winner, Wins)%>%
  distinct() %>%
  rename(Team = winner) %>%
  right_join(teams) %>%
  mutate(Wins = replace_na(Wins, 0)) %>%
  mutate(Losses = week - Wins) %>%
  mutate(WinPerc =Wins/(Wins+Losses)) %>%
  arrange(-WinPerc) %>%
  DT::datatable(escape = FALSE,options = list(lengthChange = FALSE, dom = 't')) 

```

Misc Comparisons
=====================================  
    
Column {.tabset}
-------------------------------------

### What Positions Are Predicted Most Accurately

```{r, fig.width=3.5, fig.align='center'}
WeeklyEstimates %>% as_tibble() %>% 
  filter(!is.na(ActualMinusPredicted)) %>%
  group_by(Position) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  ggplot(aes(x=ActualMinusPredicted)) +
  geom_density(aes(fill = Position)) +
  geom_histogram(aes(fill = Position)) +
  facet_wrap(~forcats::fct_reorder(Position,-n), ncol = 1) +
  geom_vline(aes(xintercept = 0, color = "red")) +
  theme(legend.position = "none", 
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  labs(title = "Histogram of How Bad (or Good)",
  subtitle = "ESPN Projects Each Position", 
       y = "", x = "Actual Points Scored Minus Predicted")
```

### Performance By Position Compared To Other Teams

```{r}
library(ggiraphExtra)

radar =
  playerdata %>%
  filter(Starter=="Starter") %>%
  filter(!is.na(Team)) %>%
  group_by(Team,Position) %>%
  summarise(Sum = sum(Actual, na.rm = TRUE)) %>%
  # filter(Position=="Running Back") %>%
  group_by(Position) %>%
  arrange(Position) %>%
  mutate(rank = rank(Sum)) 
  


radardata =
radar %>%
  ungroup() %>%
  # rename(group = Team) %>%
  # mutate_at(vars(-group,-Position), funs(scales::rescale)) %>%
  select(-Sum) %>%
  spread(Position,rank) %>%
  # bind_rows(
    # tibble(Team = "Baseline")
  # ) %>%
  replace(is.na(.), 0) %>%
  # filter(Team==params$team) %>%
  mutate(Team = str_replace_all(Team,"'",""))

radardata %>%
  ggRadar(aes(color=Team), interactive = TRUE, rescale = FALSE, alpha = .1, size = 1, legend.position = "right")

```


